---
title: "Reproducible Research: Peer Assessment 1"
author: "Akshat S Prakash"
date: "08-Mar-2020"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 2 - Course Project 1
## Peer-Graded Assignment

### 1. Code for reading in the dataset and/or processing the data
* Unzip and read the dataset
```{r reading}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activity <- read.csv('activity.csv')
activity$date<-as.Date(as.character(activity$date))
```

### 2. Histogram of the total number of steps taken each day
* Plot the histogram
```{r histogram}
library(ggplot2)
stepActivity <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)

g <- qplot(stepActivity, xlab = 'Total steps per day', ylab = 'Frequency using binwidth 1000', binwidth=1000)
plot(g)
```

### 3. Mean and median number of steps taken each day
```{r averages}
meanStep <- mean(stepActivity)
medianStep <- median(stepActivity)

g <- g + geom_vline(color= 'blue', xintercept = meanStep) +
  geom_vline(color= 'red', linetype = "dotdash", xintercept = medianStep) +
  ggtitle("Total Steps Each Day")
plot(g)
```

* Mean: `r format(meanStep, digits=2, nsmall=2)`
* Median: `r format(medianStep, digits=2, nsmall=2)`

### 4. Time series plot of the average number of steps taken
```{r timeseries}
avg_steps_taken <- aggregate(steps ~ interval, activity, mean)


g <- ggplot(data=avg_steps_taken, aes(x=interval, y=steps)) +
  geom_line() + xlab("5-minute interval") +
  ylab("Average number of steps taken") +
  ggtitle("Average Number of Steps per Day by Interval")
plot(g)

```

### 5. The 5-minute interval that, on average, contains the maximum number of steps
* Maximum number of steps is at: `r avg_steps_taken$interval[which.max(avg_steps_taken$steps)]`

### 6. Code to describe and show a strategy for imputing missing data
```{r missing}
missing <- is.na(activity$steps)
```
* The total number of missing values in the dataset is: `r sum(missing)`


### 7. Histogram of the total number of steps taken each day after missing values are imputed

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

### 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
